Meeting the Challenges of Modeling Astrophysical Thermonuclear Explosions: \castro, \maestro, and the \amrex\ Astrophysics Suite

### 2017-10-05 astronum_2017/astronum-amrex.tex
more work

GPUs

     . show Don's plot

@@

Figure~\ref{fig:maestro_scaling} shows strong scaling for \maestro\ on the 3-d
XRB problem.

show new scaling

Our latest focus has been on GPUs

@@

\ack The work at Stony Brook was supported by DOE/Office of Nuclear
Physics grant DE-FG02-87ER40317 and NSF award AST-1211563.  An award
of computer time was provided by the Innovative and Novel
Computational Impact on Theory and Experiment (INCITE) program.  This
research used resources of the Oak Ridge Leadership Computing Facility
at the Oak Ridge National Laboratory, which is supported by the Office
of Science of the U.S. Department of Energy under Contract
No.\ DE-AC05-00OR22725.  This research used resources of the National
Energy Research Scientific Computing Center, which is supported by the
Office of Science of the U.S. Department of Energy under Contract
No.\ DE-AC02-05CH11231.  Results in this paper were obtained using the
high-performance LIred computing system at the Institute for Advanced
Computational Science at Stony Brook University, which was obtained
through the Empire State Development grant NYS \#28451.  The GPU work
benefited from a grant of an Titan X Pascal board from Nvidia Corp
through the GPU Grant Program.

>>>>

A key design goal of our application codes is performance portability.
We want the same kernels to run on clusters, manycore machines (like
Intel KNL), and GPU-based machines.  Our development has balanced this
need with architecture-specific optimizations to maximize code reuse.

@@

Figure~\ref{fig:scaling} shows strong scaling for \maestro\ on the 3-d
XRB problem.  We see that the code scales well to $\mathcal{O}(10^4)$
processors.  The main limitation to the scaling at the moment are the
multigrid solves used to enforce the projection (in particular the
nodal solver).  \maestro\ also does not yet take advantage of the
tiling capabilities offered in \amrex.  These will be addressed in the
future as we port \maestro\ to the C++ \amrex\ codebase to take advantage
of ongoing development.

Castro scaling...

Our latest focus has been on GPU ports of our application codes.  A
small proxy app, \starlord, was created from \castro\ with just the
hydrodynamics and stellar equation of state.  It uses a simple
method-of-lines formulation of the hydrodynamics and advects 13
nuclear species in addition to the hydrodynamics.  To offload work on
GPUs, GPU support was added directly into \amrex.  In \amrex, grids
consist of a number of boxes that enable domain decomposition for
parallelism.  An iterator loops over all of the boxes at the same
level of refinement and passes a data pointer into a Fortran kernel
function where the work is done. For the simulation state data that
resides in each box, we have modified the memory allocator so that it
can use a CUDA allocator (either device or managed memory).  The
domain iterator is configured to handle data motion to and from the
device, so that the compute kernels can operate on data which is
presumed to already be there, and the computation is decoupled from
the memory management. Computation on the data can then be performed
with OpenACC, CUDA Fortran, or (more recently) OpenMP 4.5. We have
also built CUDA compute support into \amrex\ so that a compute kernel
can be transparently operated on using CUDA Fortran without
substantially modifying the kernels (and we anticipate using a similar
strategy to use OpenACC and/or OpenMP in the future). This helps
ensure that we can continue to maintain performance portability in our
simulation codes, by decoupling the physics algorithms from the
backend support used to implement them on various compute
architectures.\MarginPar{add CPU time for reference}

Figure~\ref{fig:scaling} shows the performance of \starlord\ on the
summitdev platform at OLCF\footnote{summitdev consists of 2 IBM Power8
  processors and 4 Nvidia Pascal GPUs per node.}.  The highest GPU
count (128) corresponds to 32 nodes on summitdev.  We see a nearly
linear speedup with the number of GPUs, indicating good weak scaling
on this machine.  Efforts are underway to complete the port of the GPU
developments into \castro.

\begin{figure}[t]
\centering
\includegraphics[width=0.48\linewidth]{CUDA-Fortran-VODE-Integrator}
\begin{minipage}[b]{0.48\linewidth}
\caption{\label{fig:scaling} (left) \maestro\ strong scaling on NERSC edison
and OCLF titan for a 3-d XRB problem.  Two different problem sizes are show.
(right) \starlord\ scaling across nodes on the OLCF summitdev platform, using
up to 4 GPUs per node.\vspace{2em}}
\end{minipage}
\end{figure}

A parallel effort is porting our microphysics, in particular reaction
networks to GPUs.  Our strategy is to do the entire ODE integration on
the GPU, i.e., the data for a patch of zones is passed to the GPU, all
righthand size and Jacobian evaluations and the timestepping itself is
done on the GPU and once all zones are burned, we access the data as
needed on the CPU.  To enable this, we ported our workhorse ODE
integrator (VODE~\cite{vode}) to CUDA Fortran. This required extensive
rewrites of the internals of VODE, which was originally written using
Fortran 77 syntax unsupported by CUDA Fortran. Accelerating VODE with
CUDA Fortran has proven successful, and we now see significant
performance gains with the CUDA version of our reaction networks, even
for moderate sized networks (a 13-isotope standard network).
Figure~\ref{fig:cudaode} shows the speed-ups on a GPU vs.\ single CPU
core.

@@

\ack The work at Stony Brook was supported by DOE/Office of Nuclear
Physics grant DE-FG02-87ER40317 and NSF award AST-1211563.  An award
of computer time was provided by the Innovative and Novel
Computational Impact on Theory and Experiment (INCITE) program.  This
research used resources of the Oak Ridge Leadership Computing Facility
at the Oak Ridge National Laboratory, which is supported by the Office
of Science of the U.S. Department of Energy under Contract
No.\ DE-AC05-00OR22725.  We appreciate the efforts of the OLCF (and in
particular Fernanda Foertter) in organizing the GPU hackathons.  This
research used resources of the National Energy Research Scientific
Computing Center, which is supported by the Office of Science of the
U.S. Department of Energy under Contract No.\ DE-AC02-05CH11231.
Results in this paper were obtained using the high-performance LIred
computing system at the Institute for Advanced Computational Science
at Stony Brook University, which was obtained through the Empire State
Development grant NYS \#28451.  The GPU work benefited from a grant of
an Titan X Pascal board from Nvidia Corp through the GPU Grant
Program.
